### 2014-10-10 coverletter/coverletter.tex
some changes

\documentclass[a4paper,10pt]{letter}
% Some of the article customisations are relevant for this class
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{hyperref}
\address{Sivaram Ambikasaran\\
Courant Institute of Mathematical Sciences\\
New York University\\
251 Mercer Street\\
New York, N.Y. 10012-1185\\
sivaram@cims.nyu.edu\\
650-996-0909}
\date{\today}
% Alternatively, these may be set on an individual basis within each letter environment.

%\makelabels % this command prints envelope labels on the final page of the document

\begin{document}

\begin{letter}{Panayot S. Vassilevski\\
Editor-in-Chief\\
Numerical Linear Algebra with Applications}

\opening{Dear Editor,} % eg Hello.

\onehalfspacing

I am submitting a manuscript titled {\em Generalized Rybicki Press algorithm}. This manuscript discusses an $\mathcal{O}(N)$, direct, numerically stable algorithm for inverting covariance matrices, whose entries are obtained as sum of exponentials. The algorithm is in-fact true for a large class of structured dense matrices known as semi-separable matrices. This manuscript also serves to formally announce the release of the implementation of the $\mathcal{O}(N)$ solver for semi-separable matrices made available at the following repository: \url{https://github.com/sivaramambikasaran/ESS}.

\closing{Regards,\\
Sivaram Ambikasaran}

\end{letter}
\end{document}