### 2015-05-02 First_Review/coverletter/coverletter.tex

\documentclass[a4paper,10pt]{letter}
% Some of the article customisations are relevant for this class
\usepackage{setspace}
\usepackage{fullpage}
\usepackage{hyperref}
\address{Sivaram Ambikasaran\\
Courant Institute of Mathematical Sciences\\
New York University\\
251 Mercer Street\\
New York, N.Y. 10012-1185\\
sivaram@cims.nyu.edu\\
650-996-0909}
\date{\today}
% Alternatively, these may be set on an individual basis within each letter environment.

%\makelabels % this command prints envelope labels on the final page of the document

\begin{document}

\begin{letter}{Panayot S. Vassilevski\\
Editor-in-Chief\\
Numerical Linear Algebra with Applications}

\opening{Dear Editor,} % eg Hello.

\onehalfspacing

I am submitting the revised manuscript titled {\em Generalized Rybicki Press algorithm}. I would like to thank you and the reviewer for careful evaluation of my manuscript. The reviewer's detailed evaluation of the paper and his comments have certainly helped me improve the paper. I am herewith sending the revised manuscript and my response to the reviewer's comments.

\closing{Regards,\\
Sivaram Ambikasaran}

\end{letter}
\end{document}